---
title: "Read OHC data"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```


```{r libraries}
library(stars)
library(tidyverse)
library(colorspace)
library(lubridate)
```

```{r set_theme}

theme_set(theme_bw())

```


# Read data

Observation-based estimates of the ocean heat content (OHC) were downloaded from:  
http://www.ocean.iap.ac.cn

Latest relevant publication:  
Cheng et al. (2022)  
https://doi.org/10.1175/JCLI-D-21-0895.1

```{r paths}

path_ohc <-
  "/nfs/kryo/work/updata/cheng/IAP_Ocean_heat_content_0_2000m/"

path_root           <-
  "/nfs/kryo/work/jenmueller/ocean_heat_and_carbon_content/data_preprocessed"

```


```{r read_ohc_data, fig.asp=1}

files_ohc <- list.files(path_ohc,
                        pattern = "OHC_")

files_ohc <- files_ohc[481:961]
dates_ohc <- as.Date(paste(str_sub(files_ohc, 22, 25), str_sub(files_ohc, 33, 34),"15", sep = "/"),
                     format = "%Y/%m/%d")

decades_ohc <- dates_ohc[seq(1,length(dates_ohc),120)]

ohc_stars <- read_stars(paste0(path_ohc,
                               files_ohc),
                        along = list(date = dates_ohc),
                        quiet = TRUE)

```

# Regridding

OHC data are originally provided:  

- with monthly resolution  
- on a 1x1 degree horizontal grid  
- integration depth of 100, 300, 700, 1500, and 2000m  
- for the period 1960 - 2021

For the purpose of this exploratory analysis, we regrid the data to:  

- decadal resolution  
- a 5x5 degree horizontal grid  
- and focus on the 1980 - 2019 period  

The regridding is done by calculating the mean values according within the coarser spatio-temporal grid.

```{r regridding}

# horizontal regridding
r5 = st_as_stars(st_bbox(ohc_stars), dx = 5, dy = 5)
ohc_stars_mean = st_warp(ohc_stars, r5)
rm(r5)

# aggregate to decades
ohc_stars_mean <- 
  aggregate(ohc_stars_mean, by = decades_ohc, FUN = mean, na.rm = TRUE)

ohc_stars_mean <- ohc_stars_mean %>%
  slice(time, 1:length(decades_ohc)-1)

# convert decade label
decades_ohc <- paste0(str_sub(st_get_dimension_values(ohc_stars_mean, "time"), 1,4), "s")
ohc_stars_mean <- st_set_dimensions(ohc_stars_mean, "time", values = decades_ohc)



```

# Total OHC column inventories

Below, we're plotting the regridded column inventories of the total OHC for the 5 integration depth limits and the four decades since 1980.

```{r absolute_column_inventories}



for (i_name in names(ohc_stars_mean)) {
  # i_name <- names(ohc_stars_mean)[1]
  
  print(
    ggplot() +
      geom_stars(data = ohc_stars_mean,
                 aes(x, y, fill = !!sym(i_name) / 1e10)) +
      facet_wrap(~ time) +
      coord_quickmap(expand = 0) +
      scale_fill_viridis_b(name = expression(OHC ~ (10 ^ 10 ~ J * m ^ -2)),
                           n.breaks = 10) +
      labs(title = paste(
        "Integration depth:", str_remove(i_name, "OHC")
      )) +
      theme(
        axis.title = element_blank(),
        legend.position = "top",
        legend.key.width = unit(2, 'cm')
      )
  )
  
}

```


# Decadal changes

Next, we're calculating the decadal difference between the OHC column inventories for the different integration depth limits.

## Full column inventories

Below, we plotted the decadal column inventory changes of the OHC, integrated from the surface to the given integration depth limit (rows in the panel plots).

```{r decadal_changes_total_column_inventories, fig.asp=1}

ohc_decade <- ohc_stars_mean %>% 
  as_tibble()

ohc_decade <- ohc_decade %>% 
  rename(lat = y,
         lon = x,
         decade = time)

ohc_decade_change <- ohc_decade %>%
  arrange(decade) %>%
  pivot_longer(
    starts_with("OHC"),
    values_to = "OHC",
    names_to = "depth",
    names_prefix = "OHC"
  ) %>% 
  mutate(depth = as.numeric(depth)) %>% 
  group_by(lon,lat,depth) %>% 
  mutate(delta_OHC = OHC - lag(OHC),
         delta_decade = paste(lag(decade), decade, sep = "-")) %>% 
  ungroup() %>% 
  filter(decade != first(decade)) 

ohc_decade_change <- ohc_decade_change %>% 
  mutate(delta_OHC = delta_OHC / 1e9)


ohc_decade_change %>%
  ggplot(aes(lon, lat, fill = delta_OHC)) +
  geom_raster() +
  coord_quickmap(expand = 0) +
  scale_fill_binned_divergingx(
    palette = "RdBu",
    rev = TRUE,
    name = expression(Delta * OHC ~ (10 ^ 9 ~ J * m ^ -2)),
    n.breaks = 10
  ) +
  facet_grid(depth ~ delta_decade) +
  theme(axis.title = element_blank(),
        legend.position = "top",
        legend.key.width = unit(2, 'cm'))

```

## Depth layer column inventories

Here, we're plotting decadal OHC changes for individual depth layers, i.e. we're subtracting the shallower column inventory (eg 700m) from each deeper level (eg 1500m) to derive the changes occuring between the layers (eg between 700 and 1500m).

### Integrated change

```{r decadal_changes_column_inventories_depth_layer_total, fig.asp=1}

ohc_decade_change_layer <- ohc_decade_change %>%
  arrange(depth) %>%
  group_by(lon, lat, delta_decade) %>%
  mutate(delta_OHC_layer = delta_OHC - lag(delta_OHC, default = 0),
         layer_thickness = depth - lag(depth, default = 0),
         delta_OHC_layer_per_m = delta_OHC_layer / layer_thickness,
         layer = paste(lag(depth, default = 0), depth, sep = "-"),
         layer = fct_inorder(layer)) %>%
  ungroup()
  
ohc_decade_change_layer %>% 
  ggplot(aes(lon, lat, fill = delta_OHC_layer)) +
  geom_raster() +
  coord_quickmap(expand = 0) +
  scale_fill_binned_divergingx(
    palette = "RdBu",
    rev = TRUE,
    name = expression(Delta * OHC ~ (10 ^ 9 ~ J * m ^ -2)),
    n.breaks = 10
  ) +
  facet_grid(layer ~ delta_decade) +
  theme(axis.title = element_blank(),
        legend.position = "top",
        legend.key.width = unit(2, 'cm'))
  
```

### Concentration change

To account for the different thickness of the depth layer, we're dividing the integrated change across the full vertical extend of the layer by it's thickness, i.e. we derive a mean OHC change of the layer in concentration units.

```{r decadal_changes_column_inventories_depth_layer_concentration, fig.asp=1}

ohc_decade_change_layer %>%
  ggplot(aes(lon, lat, fill = delta_OHC_layer_per_m * 1e3)) +
  geom_raster() +
  coord_quickmap(expand = 0) +
  scale_fill_binned_divergingx(
    palette = "RdBu",
    rev = TRUE,
    name = expression(Delta * OHC ~ (10 ^ 6 ~ J * m ^ -3)),
    n.breaks = 10
  ) +
  facet_grid(layer ~ delta_decade) +
  theme(axis.title = element_blank(),
        legend.position = "top",
        legend.key.width = unit(2, 'cm'))

```


```{r coordinate_transformation, eval=FALSE}


# st_crs(ohc_stars_mean) = 4326
# st_transform(ohc_stars_mean, st_crs(9001))

```

